Service Oriented ARC Storage; \\An Overview and Design

### 2008-12-05 doc/storage_system/ARCStorage_src/ARCStorage.tex
Final update after Alex' comments

The following test where carried out:

\begin{itemize}
\item Test 1: The distribution of data after uploading 10 large size (1 GB)
  files with one replica each.  
\item Test 2: The distribution of files after uploading 100 small size (1 kB)
  files with one replica each, and with simultanious uploading
\item Test 3: The distribution of data after repeating Tests 1 and 2 with three replicas.  
\item Test 4: System behavior while some of the shepherd becomes offline.   
\end{itemize}

All the tests were performed and worked as expected. However, some of
the tests deserve extra attention.

Fig. \ref{fig:FileDistribution} illustrate two test results,
corresponding to Test 2 and Test 3: The orange bars show the distribution after uploading 100 small files
simultaneously, without replication. This gives an indication of how
the system balances the load of many clients uploading at the same
time. The green bars show distribution with the same kind of
uploading, but with the clients requesting three replicas for each file. The
difference here is that the system simultaneously has to handle both
replication and the clients requesting uploads. When the Bartender chooses
a location for which to put a replica, it generates a list of
Shepherds that (a) do not have a replica of the file, and (b) is not
already in the process of uploading the replica. Next, it draws a
Shepherd from the list at random, using a uniform random number
generator. When uploading without asking for replicas we would
therefore expect a relatively flat distribution of the files, as can
be seen in Fig. \ref{fig:FileDistribution}.

@@

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llllll}
\hline \hline
&\small{UPPMAX}&\small{USIT-1}&\small{USIT-2}&\small{USIT-3}&\small{NIIF}\\
\hline
           \small{Load(MB)}&6683&7638&7650&4773&2289\\
\hline
\end{tabular}
\caption{Overall load distribution on the storage nodes after Tests 1,
2 and 3 were finished}
\label{table:load distribution}  
\end{table}

Table \ref{table:load distribution} shows the disk usage after Tests
1, 2 and 3. Worth noticing here is that the bandwidth NIIF was
significantly lower than between UPPMAX and USIT. If the bandwith is
saturated and the storage node is busy with, e.g., checksumming a
large file, the heartbeat from this Shepherd may be delayed, causing the
Bartender not to choose this Shepherd for the next replica. This may explain the
relatively low storage load on the  NIIF storage node.

%% \begin{table}[ht]
%% \centering
%% %% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
%% \begin{tabular}[width = \columnwidth]{llllll}
%% \hline \hline
%% &\small{UPPMAX}&\small{USIT-1}&\small{USIT-2}&\small{USIT-3}&\small{NIIF}\\
%% \hline
%%            \small{Before(MB)}&4888&7820&6843&7820&3320\\
%%            \small{During(MB)}&8798&9776& - &9775&2344 \\
%%            \small{After(MB)}&8798&6843&4888&8798&1367\\
%% \hline
%% \end{tabular}
%% \caption{Storage load before, during and after a Shepherd outage.}
%% \label{table:self healing}  
%% \end{table}

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llll}
\hline \hline
&\small{Before (MB)}&\small{During (MB)}&\small{After (MB)}\\
\hline
\small{UPPMAX}&4888&8799&8798\\
\small{USIT-1}&7820&9778&6843\\
\small{USIT-2}&6843&-&4888\\
\small{USIT-3}&7820&9774&8798\\
\small{NIIF}&3320&2344&1367\\
\hline
\small{Sum}&30691&30695&30694\\
\hline
\end{tabular}
\caption{Storage load before, during and after a Shepherd outage.}
\label{table:self healing}  
\end{table}

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking out one Shepherd from the system, and later reinsert it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished; and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted and that the storage load evens out 
again when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. We also see that the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaniously
starts uploading replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

@@

The proposed system is still in an early phase of development, but our test
results depict that the architecture is strong enough to handle
the challenges for distributed large scale storage. Much
effort is required to make the system production
ready. However, we believe that continuing in the same direction
will enable us to provide a persistent  and flexible storage system
which can fulfill the needs of the scientific community.

@@

Even though we have more work in front of us, we have shown that the
ARC Storage already is in a state where initial real-life tests can be
done. We have described a simple, yet strong architecture which we
believe will benefit communities in need of a lightweight,
yet distributed storage solution.

@@

We wish to thank Mattias Ellert for helpful discussions and guidance
on the ARC middleware. In addition, we like to thank UPPMAX, NIIF and USIT for
providing resources for running the storage tests.

>>>>

The following tests were carried out:

\begin{itemize}
\item Test 1: The distribution of data after uploading 10 large size (1 GB)
  files with one replica each.  
\item Test 2: The distribution of files after uploading 100 small size (1 kB)
  files with one replica each, and with simultanious uploading
\item Test 3: The distribution of data after repeating Tests 1 and 2 with three replicas.  
\item Test 4: System behavior while some of the shepherds are offline.   
\end{itemize}

In the tests we used two client machines: One ASUS eee 901 on a wireless network in Uppsala, 
and one Dell PowerEdge 1425SC with 10Gb ethernet connection at USIT. All the tests were performed and worked as expected: All the files were both 
uploaded and downloaded with correct checksums and they all got the requested 
number of replicas within a reasonable time. However, some of
the tests deserve extra attention.

Fig. \ref{fig:FileDistribution} illustrates two test results,
corresponding to Test 2 and Test 3: The orange (light grey) bars show the distribution after uploading 100 small files
simultaneously, without replication. This gives an indication of how
the system balances the load of many clients uploading at the same
time. The green (dark grey) bars show the distribution with the same kind of
uploading, but with the clients requesting three replicas for each file. The
difference here is that the system simultaneously has to handle both
replication and the clients requesting uploads. When the Bartender chooses
a location for which to put a replica, it generates a list of
Shepherds that (a) do not have a replica of the file, and (b) are not
already in the process of uploading the replica. Next, it draws a
Shepherd from the list at random, using a uniform random number
generator. When uploading without asking for replicas we would
therefore expect a relatively flat distribution of the files, as can
be seen in Fig. \ref{fig:FileDistribution}.

@@

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llllll}
\hline \hline
&\small{UPPMAX}&\small{USIT-1}&\small{USIT-2}&\small{USIT-3}&\small{NIIF}\\
\hline
           \small{Load(GB)}&6.683&7.638&7.650&4.773&2.289\\
\hline
\end{tabular}
\caption{Overall load distribution on the storage nodes after Tests 1,
2 and 3 were finished}
\label{table:load distribution}  
\end{table}

Table \ref{table:load distribution} shows the disk usage after Tests
1, 2 and 3. It is worth noticing here that the bandwidth to NIIF was
significantly lower than between UPPMAX and USIT. If the bandwith is
saturated and the storage node is busy, e.g., checksumming a
large file, the heartbeat from this Shepherd may be delayed, causing the
Bartender not to choose this Shepherd for the next replica. This may explain the
relatively low storage load on the  NIIF storage node.

%% \begin{table}[ht]
%% \centering
%% %% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
%% \begin{tabular}[width = \columnwidth]{llllll}
%% \hline \hline
%% &\small{UPPMAX}&\small{USIT-1}&\small{USIT-2}&\small{USIT-3}&\small{NIIF}\\
%% \hline
%%            \small{Before(GB)}&4.888&7.820&6.843&7.820&3.320\\
%%            \small{During(GB)}&8.798&9.776& - &9.775&2.344 \\
%%            \small{After(GB)}&8.798&6.843&4.888&8.798&1.367\\
%% \hline
%% \end{tabular}
%% \caption{Storage load before, during and after a Shepherd outage.}
%% \label{table:self healing}  
%% \end{table}

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llll}
\hline \hline
&\small{Before (GB)}&\small{During (GB)}&\small{After (GB)}\\
\hline
\small{UPPMAX}&4.888&8.798&8.798\\
\small{USIT-1}&7.820&9.778&6.843\\
\small{USIT-2}&6.843&-    &4.888\\
\small{USIT-3}&7.820&9.774&8.798\\
\small{NIIF}  &3.320&2.344&1.367\\
\hline
\small{Sum}&30.69&30.69&30.69\\
\hline
\end{tabular}
\caption{Storage load before, during and after a Shepherd outage.}
\label{table:self healing}  
\end{table}

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking one Shepherd out of the system, and later reinserting it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished; and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted and that the storage load evens out 
again when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. We also see that the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaniously
starts uploading replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

@@

The proposed system is still in an early phase of development, but our test
results demonstrate that the architecture is robust enough to handle
the challenges for distributed large scale storage. Much
effort is required to make the system production
ready. However, we believe that continuing in the same direction
will enable us to provide a persistent  and flexible storage system
which can fulfill the needs of even the most demanding scientific community.

@@

%% Even though we have more work in front of us, we have shown that the
%% ARC Storage already is in a state where initial real-life tests can be
%% done. We have described a simple, yet robust architecture which we
%% believe will benefit communities in need of a lightweight,
%% yet distributed storage solution.
While the work in the above mentioned areas is still in progress, we have 
already shown that the ARC Storage already is in a state where initial 
real-life tests can be done. We have described a simple, yet strong 
architecture which we believe will benefit communities in need of a lightweight, 
yet distributed storage solution.

@@

We wish to thank Mattias Ellert for helpful discussions and guidance
on the ARC middleware, to Oxana Smirnova, Alex Read and David Cameron 
for vital comments and proof reading. In addition, we like to thank UPPMAX, NIIF and USIT for
providing resources for running the storage tests.
